Miniproject - Analytic Number Theory

### 2021-12-08 Algebraic geometry/submission.tex
Added Dirichlet sum of f over primes to ANT

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Proj}{\mathbb{P}}
\newcommand{\Aff}{\mathbb{A}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Gr}{\mathrm{Gr}}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\contradiction}{\lightning}
\newcommand{\extpow}{\mathchoice{{\textstyle\bigwedge}}
    {{\bigwedge}}
    {{\textstyle\wedge}}
    {{\scriptstyle\wedge}}}
\newcommand{\vspan}{\mathrm{span}}
\newcommand{\divides}{\ | \ }
\newcommand\restr[2]{{
    \left.\kern-\nulldelimiterspace
    #1
    \vphantom{\big|}
    \right|_{#2}
}}

@@

    For the other direction, we use a quite lengthy degree argument.
    Sadly, the argument I found is quite full of technical detail.
    I try to present it as clearly as possible, at the cost of being not always completely formal. 
    But honestly, the proof contains enough indices already.

@@

    If $(a, d) = (i, v)$ then $(b, c) \neq (j, u)$ and $x_{b, c} \divides f_{i, j, u, v}$.
    Thus
    \begin{equation*}
        x_{b, c} \ x_{i, u} \ x_{j, v} \divides \mathrm{lt}(F) = \epsilon \ x_{i, u} \ x_{j, v} \ f_{i, j, u, v}
    \end{equation*}
    with $i < j < u < v$ and $i < b < c < v$. 
    No matter how $j, u, b, c$ are ordered relatively to each other, we see that in each possible case $x_{b, c} \ x_{i, u} \ x_{j, v} \in J$:
    \begin{align*}
        j < u < b < c \ &\Rightarrow \ x_{j, v} x_{b, c} \in J \\
        j < b < u < c \ &\Rightarrow \ x_{j, v} x_{b, c} \in J \\
        j < b < c < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
        b < j < u < c \ &\Rightarrow \ f_{i, b, c, v} \ x_{i, c} \ x_{b, v} \succ f_{i, j, u, v} \ x_{i, u} \ x_{j, v} \ \contradiction \\
        b < j < c < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
        b < c < j < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
    \end{align*}
    Hence $\mathrm{lt}(F) \in J$.

@@

    Finally, one must also consider the third ``big'' case, namely that
    \begin{equation*}
        \mathrm{lt}(F) = \epsilon \ x_{i, j} \ x_{u, v} \ f_{i, j, u, v} 
    \end{equation*}
    Again, you can do this similarly as before, but now two monomials ``cancel out''.
    I will not present this here either.
\end{proof}

>>>>

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Proj}{\mathbb{P}}
\newcommand{\Aff}{\mathbb{A}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Gr}{\mathrm{Gr}}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\contradiction}{\text{contradiction}}
\newcommand{\extpow}{\mathchoice{{\textstyle\bigwedge}}
    {{\bigwedge}}
    {{\textstyle\wedge}}
    {{\scriptstyle\wedge}}}
\newcommand{\vspan}{\mathrm{span}}
\newcommand{\divides}{\ | \ }
\newcommand\restr[2]{{
    \left.\kern-\nulldelimiterspace
    #1
    \vphantom{\big|}
    \right|_{#2}
}}

@@

    For the other direction, we use a quite lengthy degree argument.
    Sadly, the only argument I came up with is extremely technical.
    We try to present it as clearly as possible, at the cost of only sketching some parts.
    In the end, I checked it using Computer Algebra, and everything fits together.

@@

    If $(a, d) = (i, v)$ then $(b, c) \neq (j, u)$ and $x_{b, c} \divides f_{i, j, u, v}$.
    Thus
    \begin{equation*}
        x_{b, c} \ x_{i, u} \ x_{j, v} \divides \mathrm{lt}(F) = \epsilon \ x_{i, u} \ x_{j, v} \ f_{i, j, u, v}
    \end{equation*}
    with $i < j < u < v$ and $i < b < c < v$. 
    No matter how $j, u, b, c$ are ordered relatively to each other, we see that in each possible case $x_{b, c} \ x_{i, u} \ x_{j, v} \in J$:
    \begin{align*}
        j < u < b < c \ &\Rightarrow \ x_{j, v} x_{b, c} \in J \\
        j < b < u < c \ &\Rightarrow \ x_{j, v} x_{b, c} \in J \\
        j < b < c < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
        b < j < u < c \ &\Rightarrow \ f_{i, b, c, v} \ x_{i, c} \ x_{b, v} \succ f_{i, j, u, v} \ x_{i, u} \ x_{j, v}, \ \contradiction \\
        b < j < c < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
        b < c < j < u \ &\Rightarrow \ x_{i, u} x_{b, c} \in J \\
    \end{align*}
    Hence $\mathrm{lt}(F) \in J$.

@@

    Finally, one must also consider the third ``big'' case, namely that
    \begin{equation*}
        \mathrm{lt}(F) = \epsilon \ x_{i, j} \ x_{u, v} \ f_{i, j, u, v} 
    \end{equation*}
    Again, you can do this similarly as before, but now two monomials ``cancel out''.
    We will not present this here as well.
\end{proof}

\begin{lemma}
    Let $n \geq 4$ and
    \begin{equation*}
        J := \{ \{ A, B \} \ | \ A, B \in X, \ A \cap B = \emptyset \} \subseteq X^{(2)}
    \end{equation*}
    be a set, where
    \begin{equation*}
        X := \{ 1, ..., n \}^{(2)}
    \end{equation*}
    Then for $k \geq 3$ have
    \begin{equation*}
        |\{ A \in X^{(k)} \ | \ J \subseteq A \}| = 
    \end{equation*}
\end{lemma}
\begin{proof}
    Use induction on $n$.
    In the base case $n = 4$ have that
\end{proof}

### 2021-12-08 Analytic number theory/submission.tex
Added Dirichlet sum of f over primes to ANT

Now we want to study the series

@@

For this, we are first interested in how many primes $\equiv 1, 3 \mod 4$ there are.

@@

    Now consider the element $\alpha = \omega_q - xqP \in \mathcal{O}$. Then
    \begin{align*}
        N_{K/\Q}(\alpha) &= \prod_{\sigma: K \to \C \ \text{$\Q$-field homomorphism}} \sigma(\omega_q - xqP) \\
        &= \prod_{\sigma} (\sigma(\omega_q) - xqP) = \mathrm{MiPo}_\Q(\omega_q)(xqP) = \Phi_q(xqP) \neq 1
    \end{align*}
    as $\mathrm{MiPo}_\Q(\omega_q) = \prod_\sigma (\sigma(\omega_q) - X)$.
    Hence, $\alpha$ is not a unit in $\mathcal{O}$.
    On the other hand, $(\alpha)$ is coprime to $(p_i)$ for each $p_i$, as 
    \begin{equation*}
        \omega_q = \alpha - xqP \in (\alpha) + (p_i) \quad \text{and} \quad \omega_q \in \mathcal{O}^\units
    \end{equation*}
    By our assumption, the only prime ideals in $\mathcal{O}$ are the prime ideal factors of $(p_i)$ and $(p)$ for $p \neq p_i$.
    Thus, the prime ideal factorization of $(\alpha)$ consists only of prime ideals $(p), p \neq p_i$ and it follows that $(\alpha) = (n)$ for some integer $n \geq 2$.
    As $\omega_q$ and $xqP \in \Z$ are $\Q$-linearly independent, we see that $n \divides \omega_q$ and $n \divides xqP$.
    However, the former is a contradiction, as $\omega_q \in \mathcal{O}^\units$ is a unit and no $n \geq 2$ is a unit.
\end{proof}
The book also mentions that the general case can be proven by using L-series in algebraic number fields.
\begin{corollary}[Ex (iii)]
    There are infinitely many primes $p$ with $p \equiv 1 \mod 4$.
\end{corollary}
\begin{proof}
    This is just a special case of Prop.~\ref{prop:primes_1_mod_q}.
\end{proof}
\begin{example}[Ex (iii)]
    Using a computer, we can also study the actual frequency of prime numbers $\equiv 1, 3 \mod 4$ among e.g. the first $10^8$ integers.
    This seems to indicate that both numbers are asymptotically equal.
    For example, there are 332180 primes $\equiv 1 \mod 4$ and 332398 primes $\equiv 3 \mod 4$ smaller than $10^8$.
    To find these numbers, the following python code was used.
    \lstinputlisting[language = python]{./primes.py}
\end{example}

>>>>

Interestingly, we can also study the series

@@

using Euler Products.
    \label{ex:dirichlet_sum_over_primes}
    Have for $\Re(s) > 1$ that
        \sum_p \frac {f(p)} {p^s} = \log( L(f, s) ) + O( \zeta(2\Re(s)) )
    In particular, we see that
        \lim_{s \to 1^+} \sum_p \frac {f(p)} {p^s}
    exists.
    By Taylor series expansion, see that
        \log(1 - x) = -x + O(|x|^2)
    Hence for $\Re(s) > 1$ we can consider the logarithm of the Euler Product to find
        \log( L(f, s) ) =& \log\Bigl( \prod_p \sum_{k \geq 0} \frac {f(p^k)} {p^{ks}} \Bigr) = \log\Bigl( \prod_p \frac 1 {1 - f(p)/p^s} \Bigr) \\
        =& \sum_p -\log\Bigl( 1 - \frac {f(p)} {p^s} \Bigr) = \sum_p \frac {f(p)} {p^s} + O\Bigl( \frac {f(p)^2} {p^{2\Re(s)}} \Bigr) \\
        = & \sum_p \frac {f(p)} {p^s} + O\Bigl( \underbrace{\sum_p \frac {f(p)} {p^{2\Re(s)}}}_{\leq \zeta(2\Re(s))} \Bigr) = \sum_p \frac {f(p)} {p^s} + O(\zeta(2\Re(s)))

@@

    Now consider the element $\alpha = \omega_q - xqP \in \mathcal{O}$. Then
    \begin{align*}
        N_{K/\Q}(\alpha) &= \prod_{\sigma: K \to \C \ \text{$\Q$-field homomorphism}} \sigma(\omega_q - xqP) \\
        &= \prod_{\sigma} (\sigma(\omega_q) - xqP) = \mathrm{MiPo}_\Q(\omega_q)(xqP) = \Phi_q(xqP) \neq 1
    \end{align*}
    as $\mathrm{MiPo}_\Q(\omega_q) = \prod_\sigma (\sigma(\omega_q) - X)$.
    Hence, $\alpha$ is not a unit in $\mathcal{O}$.
    On the other hand, $(\alpha)$ is coprime to $(p_i)$ for each $p_i$, as 
    \begin{equation*}
        \omega_q = \alpha - xqP \in (\alpha) + (p_i) \quad \text{and} \quad \omega_q \in \mathcal{O}^\units
    \end{equation*}
    By our assumption, the only prime ideals in $\mathcal{O}$ are the prime ideal factors of $(p_i)$ and $(p)$ for $p \neq p_i$.
    Thus, the prime ideal factorization of $(\alpha)$ consists only of prime ideals $(p), p \neq p_i$ and it follows that $(\alpha) = (n)$ for some integer $n \geq 2$.
    As $\omega_q$ and $xqP \in \Z$ are $\Q$-linearly independent, we see that $n \divides \omega_q$ and $n \divides xqP$.
    However, the former is a contradiction, as $\omega_q \in \mathcal{O}^\units$ is a unit and no $n \geq 2$ is a unit.
\end{proof}
The book also mentions that the general case can be proven by using L-series in algebraic number fields.
\begin{corollary}[Ex (iii)]
    There are infinitely many primes $p$ with $p \equiv 1 \mod 4$.
\end{corollary}
\begin{proof}
    This is just a special case of Prop.~\ref{prop:primes_1_mod_q}.
\end{proof}
\begin{example}[Ex (iii)]
    Using a computer, we can also study the actual frequency of prime numbers $\equiv 1, 3 \mod 4$ among e.g. the first $10^8$ integers.
    This seems to indicate that both numbers are asymptotically equal, which seems natural, given the result of Example~\ref{ex:dirichlet_sum_over_primes}.
    For example, there are 332180 primes $\equiv 1 \mod 4$ and 332398 primes $\equiv 3 \mod 4$ smaller than $10^8$.
    To find these numbers, the following python code was used.
    \lstinputlisting[language = python]{./primes.py}
\end{example}
